An Integral Projection Modeling Approach to Demographic Consequences of Multiple Partner Mutualisms

### 2023-04-20 Manuscript/manuscript_draft.tex
Update manuscript_draft.tex

\ali{}{I don't love this paragraph. Help. I particularly don't like how I defined multispecies mutualisms vs pairwise}
Historically, pairwise studies, where two species engage in mutualism with only each other, were central in the field of mutualisms, it has since been shown that pairwise studies are poor predictors of the effects of multi-species mutualistic interactions\cite{Afkhami2014,Palmer2010}, where more than two species engage in mutualisms.
The differences in mutualistic outcomes are due to differences in partner quality\cite{Bascompte2019,Stanton2013,Frederickson2013,Jones2015, Ness2006}, differences in the types of benefits offered\cite{Kiers2003,Afkhami2014}, and direct interactions between partners\cite{Sun2019,Heath2009,Heath2014,Grutter2003}.
For these reasons, pairwise studies cannot be accurately used to predict the outcomes of \tom{multi-species mutualisms}{I think you need to operationally define this.}\cite{Palmer2010, Stanton2013, Chamberlain2014, Song2020}.
Studies on the effects of diverse multi-species mutualism systems are necessary to help us understand the demographic effects of partner diversity in mutualisms\cite{Bascompte2019}.

Multi-species mutualism partner diversity causes varied demographic effects on the population of the focal mutualist which can be explained by a number of mechanisms.
Partners can vary in many ways: in some cases, the quality varies leading to some true mutualist partners, \tom{some freeloaders, and even some parasites}{What is the difference between a freeloader and a parasite?} \cite{Bronstein1994,Bronstein2001a,Afkhami2014,Song2020,West2007,Frederickson2013,Jones2015}; in other cases, \tom{the function of the partner can vary}{I think I know what you mean but you should clarify for other readers what is the difference between quality and function.}, each offering a different type of reward to the focal mutualist \cite{Stanton2003}.
\tom{When there is a consistent partner hierarchy based on benefits offered, a more diverse sample of the partner community may be more likely to include the highest quality partner (the one which offers the focal mutualist the most benefit)\cite{Frederickson2013}.}{I think the content of this paragraph needs to be better organized. Describing ways in which partner species differ is conceptually very different than describing mechanisms by which partner diversity can be beneficial -- yet this paragraph seems to glom them together. The ``partners are different'' content seems to fit better in the previous paragraph.}
When the fitness of a focal mutualist with multiple partners is equal to the fitness of a focal mutualist with only the highest quality partner, as shown in \tom{Figure}{This figure is great for talks but I don't recommend it for a publication.} \ref{fig:comp-samp}b, sampling effect can explain \tom{the positive effects}{When you say \emph{the} positive effects of partner diversity, does that imply that diversity effects are always positive? I think this is the first you've said that partner diversity is beneficial.} of partner diversity\cite{Batstone2018}. 
Partners can also vary functionally, meaning each partner offers a different type of reward to the focal mutualist\cite{Stachowicz2005,Bronstein2006,Stanton2003}.
When this leads to a higher fitness for the focal mutualist than any single interaction,as shown in Figure \ref{fig:comp-samp}a, complementarity explains the benefits of partner diversity\cite{Batstone2018}. 
Partners can also have asynchronous responses to the environment, either spatially\cite{Ollerton2006} or temporally\cite{Alarcon2008}.
Multiple partners can act as a 'portfolio' leading to more even fitness benefits across environmental or temporal heterogeneity, as shown in Figure \ref{fig:portfolio}, through the portfolio effect\cite{Batstone2018,Lazaro2022}.

\begin{figure}[h]
	\includegraphics[width=0.58\linewidth]{complementarity.png}
	\includegraphics[width=0.40\linewidth]{sampling_effect.png}
	\caption{asdasdf}
	\label{fig:comp-samp}
\end{figure}
\begin{figure}[h]
	\includegraphics[width=0.60\linewidth]{portfolio_effect.png}
	\caption{asdasdf}
	\label{fig:portfolio}
\end{figure}

When multiple partners are present, \tom{partner turnover}{I suggest introducing this idea a little differently. Partner diversity can have different effects depending on whether it manifests all at once or sequentially. So this leads to the question of turnover.} is also important to consider, \tom{as partner identity has been shown to impact benefits in mutualisms}{You have made this point several times.}\cite{Djieto-Lordon2005, Ness2006, Bruna2014}. 
Turnover can happen at different timescales\cite{Oliveira1999,Horvitz1986} from minutes to years. 
The degree and order of partner turnover can have important impacts on the fitness of the focal mutualist.
High degrees of partner turnover, \tom{also known as opportunism}{I've never seen partner turnover referred to as ``opportunism''.}, may offer \tom{buffering in the face of environmental variation}{This sounds a lot like portfolio effect, so I think you need to clarify how it relates.}\cite{Hegland2009}, while low degrees of partner turnover, also known as partner fidelity, \tom{can increase the benefits}{This makes it sounds like diversity is harmful - since low turnover means low diversity. Is that an example of negative effects of partner diversity?} received in a mutualism\cite{Sachs2004}.
\tom{The direction of turnover can also have a significant impact, particularly if the most beneficial partner changes across the ontogeny of the focal mutualist's life\cite{Fonseca2003}.
For example, susceptibility to enemies can change across life stages\cite{Boege2005,Barton2010}, so can defense traits, indicating that the focal mutualist benefits the most when defensive partners align with focal mutualist development\cite{Djieto-Lordon2005}.}{This is an excellent point. The second sentence here is a little clunky - see if you can fine-tune it.}

\tom{Defensive}{Better transition from concepts to ants?} ant-plant mutualisms are common interactions in which plants provide \tom{extra-floral nectar}{I would just say ``food'' -- see Beltian bodies as another food resource} and/or ``housing'' to ants which in turn defend them from herbivores\cite{Bronstein1998, Bronstein2006}. 
\tom{Ants are highly ubiquitous\cite{Schultheiss2022} and vary significantly in abundances\cite{Schultheiss2022}}{This seems like too general/vague a statement to be useful here.}, types of rewards offered\cite{Beattie1985}, and \tom{quality of interaction}{What do you mean by this and how does it differ from types of rewards?} with their plant partners\cite{Rico-Gray1989,Palmer2010}, \tom{making these interactions great systems to study the effects of partner diversity}{A simpler and more to-the-point observation is that these interactions are typically generalized: most EFN plants have multiple ant partners} within mutualisms. 
\tom{While these interactions have been well studied\cite{Ness2006,Beattie1985,Schultheiss2022} \st{in the literature}, few have considered how diversity within ant defender guilds\cite{Stanton2013} or temporal fluctuations in partner interactions\cite{Trøjelsgaard2015} affect the overall benefits of mutualism for the plant partner.}{Good sentence.}

This study focuses on the tree cholla cactus, a long lived \textit{Cylindriopuntia imbricata}, \st{long-lived} EFN-bearing plant that associates with multiple species of ant partners.
Insect herbivory negatively affects the growth of cacti, in turn reducing population growth rates\cite{Miller2009}. 
\st{The} ant partners defend the cacti from these insect herbivores, reducing the negative effects of herbivory\cite{Miller2007}. 
The multiple ant partners do not co-occur on \tom{individual}{} plants, but a single cactus may interact with multiple species over its lifetime.
Some of the ant partners are more effective defenders than others, with \tom{\textit{Liom.}}{Notice that you have not yet introduced the species, let alone the abbreviation of its latin name, so a reader will be very confused here. Try to anticipate what your readers need to stay with you.} tended reproducing plants experiencing the lowest level of herbivory as shown in \tom{Figure \ref{fig:herb}}{You have put a result in the Introduction. Sometimes this is appropriate but I don't think it is here. You have not described any methods for where this result comes from, so it is not reproducible. Make this a result.}.
\tom{This shows that although all the ants associated with the tree cholla are within one guild, they are not all equal and may have different demographic effects on the cacti.}{I think you can make this point without needing the figure, based on what we have already published from this system. It is also worth pointig out that one of the defenders was suggested to be net-negative because it deterred pollinators (Ohm and Miller 2014) but that study did not integrate demographic effects over the life cycle, which is what you do here and worth emphasizing as a source of novelty.}

@@

In this \tom{paper}{Refer to this as a study and not a paper.} we will answer three questions about the demographic effects of partner diversity in multi-species mutualisms:
\begin{enumerate}
	\item{What are the contrasting demographic effects of multiple partners?}
	\item{How do \tom{size}{size of what?} and partner identity impact the directionality of turnover between multiple partners?}
	\item{\tom{What mechanisms explain the benefits}{This question pre-supposes that there are benefits of diversity.} of partner diversity in a multispecies mutualism?}
\end{enumerate}

\tom{To answer these questions, we used a long-term dataset of demographic information (size, survival, number of flowers produced or aborted, etc.) and ant partner data (species, number of ants) to track the structure of the population across time as well as individual level impacts of ant partners on the cacti.}{I think this sentence can be sharpened. I don't think that details like flower abortion and ant number are important to include here, but I also don't think this sentence really communicates the nature or volume of the data,} 
We analyzed the \tom{impacts of partner diversity}{The vital rate models do not reflect the impact of partner diversity, they are simply vital rate estimates fit with ant category as a factor -- be careful with how you present this.} on each individual vital rate (survival, growth, floral viability, etc.) with \tom{Bayesian models}{``Bayesian models'' does not carry much meaning. A model cannot be intrinsically Bayesian. The model is the model, and Bayesian analysis is one way to estimate it's parameters.}, which we used to parameterize an Integral Projection Model (IPM).
In previous studies the effects of multiple partners, and the dynamics of their turnovers, have been identified and quantified\cite{Palmer2010, Nishi2013, Hussa2013}, but \tom{none have identified the mechanisms that explain the benefits of partner diversity}{Not sure I agree here. How are you ``identifying mechanisms'' in a way that Palmer et al. did not, for example?}.

>>>>

\ali{}{I don't love this paragraph. Help. I particularly don't like how I defined multispecies mutualisms vs pairwise}
Historically, pairwise studies, where two species engage in mutualism with only each other, were central in the field of mutualisms, it has since been shown that pairwise studies are poor predictors of the effects of multi-species mutualistic interactions\cite{Afkhami2014,Palmer2010}, where more than two species engage in mutualisms.
The differences in mutualistic outcomes are due to differences in partner quality\cite{Bascompte2019,Stanton2013,Frederickson2013,Jones2015, Ness2006}, differences in the types of benefits offered\cite{Kiers2003,Afkhami2014}, and direct interactions between partners\cite{Sun2019,Heath2009,Heath2014,Grutter2003}.
For these reasons, pairwise studies cannot be accurately used to predict the outcomes of multi-species mutualisms\cite{Palmer2010, Stanton2013, Chamberlain2014, Song2020}.
Studies on the effects of diverse multi-species mutualism systems are necessary to help us understand the demographic effects of partner diversity in mutualisms\cite{Bascompte2019}. 
Multi-species mutualism partner diversity causes varied demographic effects on the population of the focal mutualist which can be explained by a number of mechanisms.
\ali{In some cases, the quality of the benefits offered varies leading to some partners which benefit the individual, some which have neither a positive or negative impact on the individual, and some which negatively impact the host}{Is this the best way to introduce this or should I talk about quality of benefits compared to cost of rewards? AKA some partners offer a net benefit when the benefits outweigh the cost of rewards to the partner ... etc.} \cite{Bronstein1994,Bronstein2001a,Afkhami2014,Song2020,West2007,Frederickson2013,Jones2015}. In cases where the partner offers benefits, \ali{the function, or type of benefits, offered by the partner can vary}{Is this enough detail or do I need to go into examples?} \cite{Stanton2003}.

\begin{figure}[h]
	\includegraphics[width=0.58\linewidth]{Complementarity copy.png}
	\includegraphics[width=0.40\linewidth]{Sampling_Effect copy.png}
	\caption{asdasdf}
	\label{fig:comp-samp}
\end{figure}
\begin{figure}[h]
	\includegraphics[width=0.60\linewidth]{Portfolio_Effect copy.png}
	\caption{asdasdf}
	\label{fig:portfolio}
\end{figure}

This variation in partners can lead to \ali{net benefits, no net benefits, or net costs for the focal mutualist}{Is this the best way to start this paragraph? If so should I include a figure panel?}.
In this study I focus on the positive effects of partner diversity which can be explained by several mechanisms described below. 
When there is a consistent partner hierarchy based on benefits offered, a more diverse sample of the partner community may be more likely to include the highest quality partner (the one which offers the focal mutualist the most benefit)\cite{Frederickson2013}.
When this leads to the fitness of a focal mutualist interacting with multiple partners being equal to the fitness of a focal mutualist interacting with only the highest quality partner, as shown in Figure \ref{fig:comp-samp}b, sampling effect can explain the positive effects of partner diversity\cite{Batstone2018}. 
Partners can also vary functionally, meaning each partner offers a different type of reward to the focal mutualist\cite{Stachowicz2005,Bronstein2006,Stanton2003}.
When this leads to a higher fitness for the focal mutualist than any single interaction,as shown in Figure \ref{fig:comp-samp}a, complementarity explains the benefits of partner diversity\cite{Batstone2018}. 
Partners can also have asynchronous responses to the environment, either spatially\cite{Ollerton2006} or temporally\cite{Alarcon2008}.
Multiple partners can act as a 'portfolio' leading to more even fitness benefits across environmental or temporal heterogeneity, as shown in Figure \ref{fig:portfolio}, through the portfolio effect\cite{Batstone2018,Lazaro2022}.

Partner diversity can have different effects depending on whether partners are present all at once or sequentially (partner turnover) \cite{Djieto-Lordon2005, Ness2006, Bruna2014}. 
Turnover can happen at different timescales\cite{Oliveira1999,Horvitz1986} from minutes to years. 
The degree and order of partner turnover can have important impacts on the fitness of the focal mutualist.
\ali{The degree, or rapidity, of partner turnover can impact the level of benefits recieved by the focal mutualist particularly if the benefits continue to accumulate or if they saturate over time \cite{Sachs2004}.}{Not quite sure what to say here. } 
The direction of turnover can also have a significant impact, particularly if the most beneficial partner changes across the ontogeny of the focal mutualist's life\cite{Fonseca2003}.
For example, susceptibility to enemies can change across life stages\cite{Boege2005,Barton2010} so the focal mutualist benefits the most when defensive partners align with the more vulnerable life stages of focal mutualists \cite{Djieto-Lordon2005}.

\ali{Ant visitation to extra-floral nectar producing plants are a classic and well-studied example of interspecific mutualisms which are often multispecies and have dynamic turnover patterns}{... Is this a better transition? I tried to tie it in a bit more conceptually.}, making these interactions great systems to study the effects of partner diversity.
A subset of these interactions, defensive ant-plant mutualisms are common interactions in which plants provide food and/or ``housing'' to ants which in turn defend them from herbivores\cite{Bronstein1998, Bronstein2006}. 
While these interactions have been well studied\cite{Ness2006,Beattie1985,Schultheiss2022}, few have considered how diversity within ant defender guilds\cite{Stanton2013} or temporal fluctuations in partner interactions\cite{Trøjelsgaard2015} affect the overall benefits of mutualism for the plant partner.

This study focuses on the tree cholla cactus, a long lived \textit{Cylindriopuntia imbricata}, EFN-bearing plant that associates with multiple species of ant partners.
Insect herbivory negatively affects the growth of cacti, in turn reducing population growth rates\cite{Miller2009}. 
Ant partners defend the cacti from these insect herbivores, reducing the negative effects of herbivory\cite{Miller2007}. 
The multiple ant partners do not co-occur on individual plants, but a single cactus may interact with multiple types of partners over its lifetime.
While the quality of ant partners in this system has been studied, with some seen as superior defenders and some  being viewed as net-negative because they can deter pollinators \cite{Ohm2014}, no one has integrated demographic effects over the life cycle of these cacti.

@@

In this study we will answer three questions about the demographic effects of partner diversity across host lifetimes in multi-species mutualisms:
\begin{enumerate}
	\item{What are the contrasting demographic effects of multiple partners?}
	\item{How do host size and partner identity impact the directionality of turnover between multiple partners?}
	\item{What mechanisms explain the effects of partner diversity in a multispecies mutualism?}
\end{enumerate}

To answer these questions, we used a 20-year long-term dataset of individual-level demographic information and ant partner identity and quantity data to track the structure of the population across time as well as individual level impacts of ant partners on the cacti.
We analyzed the correlations between each individual vital rate (survival, growth, floral viability, etc.) and ant partners with Bayesian analysis of heirarchical models, which we used to parameterize an Integral Projection Model (IPM).
